---
title: "Exercises"
author: "lisanechyporenko"
date: "2021-02-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Exercises

7.\ Use the hfi dataset from the openintro package

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(broom)

data(hfi)
```

```{r}
hfi_2016 <- hfi %>% 
  filter(year == 2016)
```

8.\ Fit a new model that uses  `ef_money` or monetary measure to predict `hf_score`.  What does the slope tell us in the context of the relationship between human freedom and the economy in the country? 

```{r }

p <- hfi_2016 %>% ggplot(aes(ef_money,hf_score)) 
p + geom_point() + geom_smooth(method = "lm", se = FALSE)

m1 <- lm(hf_score ~ ef_money, data = hfi_2016)
tidy(m1)

```
The slope indicates that that human freedom is highly associated with a better economy.


5\. Again add region to the model from Q4. Compare the slope and $R^2$ with the model from Q4.

```{r message=FALSE, warning=FALSE}

m1_reg <- lm(hf_score ~ ef_money + region, data = hfi_2016)
tidy(m1_reg)

glance(m1)
glance(m1_reg)

```
The slope decreases from 0.504 to 0.364, but  $R^2$ increases from 0.425 to 0.744. Region is therefore a confounder for economy.


6\. Finally fit a model with `ef_money` and `pf_expression_control` as exposures and `hf_score` as outcome.  Compare the slope and $R^2$ from the models from Q1. Could `ef_money` be a confounder? 

```{r}

m2 <- lm(hf_score ~ ef_money + pf_expression_control, data = hfi_2016)
tidy(m2)
glance(m2)

```
The slope stays quite similar (0.37 without and 0.32 with ef_money), but $R^2$ increases (from 0.62 to 0.774). That means that `ef_money` is might be a confounder.



7\. Use a linear regression model (and scatter plot) with  `ef_money` as exposure and `pf_expression_control` as outcome, to study whether `ef_money` has an association with `pf_expression_control` as well. This might validate our finding that `ef_money` is a confounder between  `pf_expression_control` as exposure and 
`hf_score` as outcome from Q6.

```{r}

p <- hfi_2016 %>% ggplot(aes(ef_money,pf_expression_control)) 
p + geom_point() + geom_smooth(method = "lm", se = FALSE)

m3 <- lm(pf_expression_control ~ ef_money, data = hfi_2016)

tidy(m3)

glance(m3)

```

Based on the low $R^2$ of 0.143 it is now clear that `ef_money` and `pf_expression_control` are ot associated, therefore it is not a confounder.

